---
title: "Visualization 2"
output: html_document
---

# Setup

```{r setup}
library(tidyverse)
library(ggridges)
```

We’ll be working with NOAA weather data, which is downloaded using `rnoaa::meteo_pull_monitors` function in the code chunk below; similar code underlies the weather dataset used elsewhere in the course. Because this process can take some time, I’ll cache the code chunk.

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

# can use weather_df |> view() in the command line to see the full dataset
```

# Same Plot From Last Time
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature Plot",
    x = "Min daily temp (Degrees Celsius)",
    y = "Max daily temp (Degrees Celsius)",
    color = "Location",
    caption = "Max vs. min daily temp in 3 locations; data from RNOAA"
  )
```

# Adding Scales
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature Plot",
    x = "Min daily temp",
    y = "Max daily temp",
    color = "Location",
    caption = "Max vs. min daily temp in 3 locations; data from RNOAA"
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0", "15 C")
    ) +
  scale_y_continuous(
    position = "right",
    trans = "sqrt",
    limits = c(0, 30)
  )
```

Note that `limits` doesn't automatically update the scale/labeling on the corresponding axis, so might need to separately make adjustments to those after limiting.

# Colors
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature Plot",
    x = "Min daily temp",
    y = "Max daily temp",
    color = "Location",
    caption = "Max vs. min daily temp in 3 locations; data from RNOAA"
  ) +
  scale_color_hue(h = c(150, 300))
```

Other option: viridis::scale_color_viridis(discrete = TRUE). The color schemes from the viridis package are better for color-blindness accessibility and clearer when printing in gray-scale. But, do need to specify discrete vs. continuous with functions from this package.

# Themes
```{r}

```


